\center \fontsize{18}{20} \selectfont Revealing the Landscape of Globally Color-Dual Multi-loop Integrands

### 2023-09-28 notes/2loopNLSM.tex
Revised comments

In addition to the tree-level construction, we found that kinetic mixing
between the two gauge fields permits globally color-dual one-loop
integrands. After plugging in suitable on-shell states, the integrands
produced by semi-abelian YM correspond to those of SDYM in \cref{SDYMnGon} 
and NLSM in \cref{NLSMnGon}.  \jm{Mention CS?  Maybe something like `SDYM in (), NLSM in (), and Chern-Simons theory in \cref{sec:CSTheory}.'?}  Due to the
kinetic mixing that stems from the broken Bose symmetry, \cref{2loopTension} shows why it is
impossible to construct a two-loop integrand for semi-abelian YM theory. In \cref{2loopBoot}, we overcame this obstruction by successfully constructing two-loop NLSM numerators with an ansatz-based bootstrap. However, we found that the same approach fails for two-loop pure Yang-Mills, and pin-pointed the failure to a particular unitarity cut captured by \cref{chefsKiss}. Thus, obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry, and in this paper we have argued that off-shell locality\footnote{An example of this can already be found in the literature, where locality is relaxed in Ref.~\cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands with on-shell 4D states.} must
be abandoned for color-dual constructions of multi-loop Yang-Mills. To expand the available function space, we propose that one should consider including rational functions of kinematics, rather than just polynomials.

@@

The correspondence between CS and semi-YM can be made more precise at the level of the Feynman rules.
The color-stripped three-point vertex and propagator can be read off from the CS action in \cref{eq:CSAction} producing
\be\label{CSFrule}
\cubic{hgrey0}{nhpRed}{}{nhpRed}{}{nhpRed}{} = \epsilon^{\mu\nu\rho} \varepsilon^{\mu}_1\varepsilon^{\nu}_2\varepsilon^{\rho}_3 \qquad \qquad \propFrule{$A^\nu$}{$A^\mu$}{$k \rightarrow $}{nhpRed} = \frac{i}{k^2} \epsilon^{\mu\nu\rho} k^\rho \, .
\ee
At first glance, this seems far removed from the semi-abelian Yang-Mills analogs in \cref{semiYMvert} and \cref{semiYMprop}. After all, semi-YM makes no reference to an antisymmetric 3-tensor and the mass dimensions of the vertices do not agree. However, to construct the full on-shell amplitudes of CS theory, we must plug in the Fourier transforms of the on-shell currents, $\bar{\varepsilon}(p)$, as described in Ref.~\cite{Ben-Shahar:2021zww}. \jm{How about we use $\varepsilon_\mu$ and $\epsilon_\mu$ like Henrik and Maor?  I'm worried that the bar is getting overused a little.  Or how about $\varepsilon$ and $\tilde{\varepsilon}$ but this will eventually involve $\tilde{\bar{\varepsilon}}$ which is a little gross?} These are related to the polarizations, $\varepsilon_i$, of the Feynman rules above as follows:
\be
\varepsilon^{\mu}(p) = \epsilon^{\mu\nu\rho} p^\nu \bar{\varepsilon}^\rho(p)\, .
\ee
With this definition of the on-shell states, it is clear how the Feynman rules of \cref{CSFrule} are related to those of \cref{semiYMvert} and \cref{semiYMprop} -- we must simply canonicalize the CS kinetic term by absorbing $ \epsilon^{\mu\nu\rho} k^\rho$ into the definition of the three-point vertex:
\be
 \propFrule{$A^\nu$}{$A^\mu$}{$k \rightarrow $}{nhpRed} = \frac{i}{k^2} \epsilon^{\mu\nu\rho} k^\rho
\quad \longrightarrow \quad  \propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}\,.
\ee
This has the effect of orienting the three-point vertex of \cref{CSFrule} by making the replacement $\varepsilon_i \rightarrow \epsilon^{\mu\nu\rho} k_i^\nu \bar{\varepsilon}^\rho_i$ on one of the legs,
\begin{align}
\cubic{hgrey0}{nhpRed}{}{nhpRed}{}{nhpRed}{} =  \epsilon^{\mu\nu\rho} \varepsilon^{\mu}_1\varepsilon^{\nu}_2 (\epsilon^{\rho \alpha \beta} k_3^\alpha \bar{\varepsilon}^\beta_3) =    (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)\, ,
\end{align}
where we have used the identity $\epsilon^{\mu\nu\rho}\epsilon^{\rho\alpha\beta} = \eta^{\mu \alpha} \eta^{\nu \beta} -  \eta^{\mu \beta} \eta^{\nu \alpha}$. As desired, the redefined CS vertex is precisely the three-point vertex of semi-YM in \cref{semiYMvert}. Clearly the $D$-dimensional amplitudes of semi-YM can be mapped to the 3D observables of CS theory by making the following replacement on all of the external states of semi-abelian YM:
\be
\varepsilon^\mu_i \rightarrow \epsilon^{\mu\nu\rho} k^\nu_i \bar{\varepsilon}^\rho_i \qquad\qquad  \bar{\varepsilon}_i^\mu \rightarrow \bar{\varepsilon}^\mu_i \, .
\ee
Upon making this replacement, the amplitudes of $D$-dimensional semi-YM are mapped onto 3D CS theory, $\mathcal{A}^{\text{semi-YM}}\rightarrow \mathcal{A}^{\text{CS}}$. While this may seem like a marginal gain at tree-level, by using our $D$-dimensional construction at one-loop, the manifestly color-dual integrands of \cref{SDYMnGon} can resolve regularization ambiguity of CS theory due to its explicit dependence on $D=3$. For more background on CS theory, we refer the reader to Ref.~\cite{Ben-Shahar:2021zww}.

\jm{Do we want to mention something about how it's not obvious in this formulation of CS that the amplitudes vanish?  This also happens for NLSM where it's not obvious why odd multiplicity semi-YM amplitudes with the NLSM external states vanish.}

>>>>

In addition to the tree-level construction, we found that kinetic mixing
between the two gauge fields permits globally color-dual one-loop
integrands. After plugging in suitable on-shell states, the integrands
produced by semi-abelian YM correspond to those of SDYM in \cref{SDYMnGon} 
and NLSM in \cref{NLSMnGon}.  \jm{Mention CS?  Maybe something like `SDYM in \cref{SDYMnGon} 
and NLSM in \cref{NLSMnGon}, where Chern-Simons theory is discussed in \cref{sec:CSTheory}.'?}  Due to the
kinetic mixing that stems from the broken Bose symmetry, \cref{2loopTension} shows why it is
impossible to construct a two-loop integrand for semi-abelian YM theory. In \cref{2loopBoot}, we overcame this obstruction by successfully constructing two-loop NLSM numerators with an ansatz-based bootstrap. However, we found that the same approach fails for two-loop pure Yang-Mills, and pin-pointed the failure to a particular unitarity cut captured by \cref{chefsKiss}. Thus, obtaining a globally color-dual
integrand beyond one-loop then requires sacrificing more than just Bose
symmetry, and in this paper we have argued that off-shell locality\footnote{An example of this can already be found in the literature, where locality is relaxed in Ref.~\cite{Mogull:2015adi} in
  order to obtain color-dual two-loop integrands with on-shell 4D states.} must
be abandoned for color-dual constructions of multi-loop Yang-Mills. To expand the available function space, we propose that one should consider including rational functions of kinematics, rather than just polynomials.

@@

The correspondence between CS and semi-YM can be made more precise at the level of the Feynman rules.
The color-stripped three-point vertex and propagator can be read off from the CS action in \cref{eq:CSAction} producing
\be\label{CSFrule}
\cubic{hgrey0}{nhpRed}{}{nhpRed}{}{nhpRed}{} = \epsilon^{\mu\nu\rho} \varepsilon^{\mu}_1\varepsilon^{\nu}_2\varepsilon^{\rho}_3 \qquad \qquad \propFrule{$A^\nu$}{$A^\mu$}{$k \rightarrow $}{nhpRed} = \frac{i}{k^2} \epsilon^{\mu\nu\rho} k^\rho \, .
\ee
At first glance, this seems far removed from the semi-abelian Yang-Mills analogs in \cref{semiYMvert} and \cref{semiYMprop}. After all, semi-YM makes no reference to an antisymmetric 3-tensor and the mass dimensions of the vertices do not agree. However, to construct the full on-shell amplitudes of CS theory, we must plug in the Fourier transforms of the on-shell currents, $\bar{\varepsilon}(p)$, as described in Ref.~\cite{Ben-Shahar:2021zww}. \jm{I'm worried that the bar over epsilons is getting overused.  Howe about we use $e_\mu$ instead of $\bar{\varepsilon}_\mu$?  We could use a tilde or $\epsilon$ but those both have issues.} These are related to the polarizations, $\varepsilon_i$, of the Feynman rules above as follows:
\be
\varepsilon^{\mu}(p) = \epsilon^{\mu\nu\rho} p^\nu \bar{\varepsilon}^\rho(p)\, .
\ee
With this definition of the on-shell states, it is clear how the Feynman rules of \cref{CSFrule} are related to those of \cref{semiYMvert} and \cref{semiYMprop} -- we must simply canonicalize the CS kinetic term by absorbing $ \epsilon^{\mu\nu\rho} k^\rho$ into the definition of the three-point vertex:
\be
 \propFrule{$A^\nu$}{$A^\mu$}{$k \rightarrow $}{nhpRed} = \frac{i}{k^2} \epsilon^{\mu\nu\rho} k^\rho
\quad \longrightarrow \quad  \propFrule{$A^\nu$}{$\bar{A}^\mu$}{$k \rightarrow $}{fermion2} = \frac{i}{k^2} \eta^{\mu\nu}\,.
\ee
This has the effect of orienting the three-point vertex of \cref{CSFrule} by making the replacement $\varepsilon_i \rightarrow \epsilon^{\mu\nu\rho} k_i^\nu \bar{\varepsilon}^\rho_i$ on one of the legs,
\begin{align}
\cubic{hgrey0}{nhpRed}{}{nhpRed}{}{nhpRed}{} =  \epsilon^{\mu\nu\rho} \varepsilon^{\mu}_1\varepsilon^{\nu}_2 (\epsilon^{\rho \alpha \beta} k_3^\alpha \bar{\varepsilon}^\beta_3) =    (\varepsilon_1\bar{\varepsilon}_3)(\varepsilon_2 k_3)
  -  (\varepsilon_2\bar{\varepsilon}_3)(\varepsilon_1 k_3)\, ,
\end{align}
where we have used the identity $\epsilon^{\mu\nu\rho}\epsilon^{\rho\alpha\beta} = \eta^{\mu \alpha} \eta^{\nu \beta} -  \eta^{\mu \beta} \eta^{\nu \alpha}$. As desired, the redefined CS vertex is precisely the three-point vertex of semi-YM in \cref{semiYMvert}. Clearly the $D$-dimensional amplitudes of semi-YM can be mapped to the 3D observables of CS theory by making the following replacement on all of the external states of semi-abelian YM:
\be
\varepsilon^\mu_i \rightarrow \epsilon^{\mu\nu\rho} k^\nu_i \bar{\varepsilon}^\rho_i \qquad\qquad  \bar{\varepsilon}_i^\mu \rightarrow \bar{\varepsilon}^\mu_i \, .
\ee
Upon making this replacement, the amplitudes of $D$-dimensional semi-YM are mapped onto 3D CS theory, $\mathcal{A}^{\text{semi-YM}}\rightarrow \mathcal{A}^{\text{CS}}$. While this may seem like a marginal gain at tree-level, by using our $D$-dimensional construction at one-loop, the manifestly color-dual integrands of \cref{SDYMnGon} can resolve regularization ambiguity of CS theory due to its explicit dependence on $D=3$. For more background on CS theory, we refer the reader to Ref.~\cite{Ben-Shahar:2021zww}.

\jm{Can we mention somewhere that CS amplitudes vanish because the theory is topological?  From the point of view of semi-YM it isn't obvious why the CS amplitudes should vanish (at least it isn't obvious to me), but a similar thing happens for odd multiplicity NLSM amplitudes.}
